---
title: "The Data Cleaning Document"
format: html
---

# Onboarding

Hi there! Welcome to the group! This is "The Data Cleaning Document". Here's the process

1. Read the data in
2. Go through various ad hoc tests
3. Make sure the data is "clean"
4. Produce a plot of the **proportion of people educated** in each age group in each state
5. Produce a box plot of **proportion of people educated** for each state.
6. Produce a table of The 5 number summary (min, 1st quantile, median, 3rd quantile, max)  of **proportion of people educated** for each state.
7. Produce a table stating the 5 number summary of **proportion educated** in each state

When we get another dataset, we make a new document to clean that data. The process is:

1. Copy this document
2. Update the data
3. Check it is "clean"
4. Make the same plots and tables

It might seem repetitive, and there's often little sneaky bits in there, but it's just part of the process! 

Enjoy!

# Data Cleaning

```{r}
#| label: read-data
library(tidyverse)

education_2014 <- read_csv("raw_education_2014.csv")

education_2014

```

# Check age groups

Check to see how people have done the separations between ages

```{r}
#| label: table-ages
table(education_2014$age_group)
```

We need to make sure that they are separated by `_` only:

```{r}
#| label: fix-age-groups
fixed_age_group_education_2014 <- education_2014 |>
  mutate(
    age_group = str_replace_all(
      age_group,
      pattern = "---|--|-",
      replacement = "_"
    ),
    age_group = as.factor(age_group)
  )
```


Check if this is fixed

```{r}
#| label: check-age-groups
table(fixed_age_group_education_2014$age_group)
```

Check to see if year is right

```{r}
#| label: check-year
table(fixed_age_group_education_2014$year)
```

Good!

# Produce a plot of the **proportion of people educated** in each age group in each state

```{r}
#| label: plot-prop-age
# now let's explore the number of people studying in each age group in each state
ggplot(
  fixed_age_group_education_2014,
  aes(
    x = prop_studying,
    y = age_group
  )
) +
  geom_col() +
  facet_wrap(~state_territory, ncol = 2)

```

Why are there negative proportions?

```{r}
#| label: summarise-prop-studying
summary(fixed_age_group_education_2014$prop_studying)
hist(fixed_age_group_education_2014$prop_studying)
```

Maybe -99 means missing!? Are there other instances of -99?


```{r}
#| label: scan-missings
library(naniar)

miss_scan_count(
  data = fixed_age_group_education_2014,
  search = -99
)
```

Replace -99 with NA

```{r}
#| label: tidy-nas
tidy_age_state_education_2014 <- fixed_age_group_education_2014 |>
  mutate(
    prop_studying = na_if(prop_studying, -99)
  )

n_miss(tidy_age_state_education_2014)
```

# Produce a box plot of **proportion of people educated** for each state.

```{r}
#| label: boxplot-prop-studying
ggplot(
  tidy_age_state_education_2014,
  aes(
    x = prop_studying,
    y = state_territory
  )
) +
  geom_boxplot()
```

# Produce a table of The 5 number summary (min, 1st quantile, median, 3rd quantile, max)  of **proportion of people educated** for each state.

```{r}
#| label: table-prop-studying
#| error: true
tidy_age_state_education_2014 |>
  group_by(state_territory) |>
  summarise(
    min_study = min(prop_studying),
    q1_studying = quantile(prop_studying, probs = 0.25),
    median_study = median(prop_studying),
    q3_studying = quantile(prop_studying, probs = 0.75),
    sd_study = sd(prop_studying)
  ) |> 
  knitr::kable(digits = 3)
```

Drop NA values

```{r}
#| label: table-prop-studying-na
tidy_age_state_education_2014 |>
  group_by(state_territory) |>
  summarise(
    min_study = min(prop_studying, na.rm = TRUE),
    q1_studying = quantile(prop_studying, probs = 0.25, na.rm = TRUE),
    median_study = median(prop_studying, na.rm = TRUE),
    q3_studying = quantile(prop_studying, probs = 0.75, na.rm = TRUE),
    sd_study = sd(prop_studying, na.rm = TRUE)
  ) |> 
  knitr::kable(digits = 3)
```
