---
title: "The Data Cleaning Document"
format: html
---

# Onboarding

Hi there! Welcome to the group! This is "The Data Cleaning Document". Here's the process

1. Read the data in
2. Go through various ad hoc tests
3. Make sure the data is "clean"
4. Repeat for subsequent years of data
5. Produce a plot of the number of people studying in each age group in each state
6. Produce a box plot, and a table that shows The 5 number summary (min, 1st quartile, median, 3rd quartile, max) of number educated for each state.
7. Produce a table stating the summary of proportion educated in each state

When we get another dataset, we make a new document to clean that data. The process is:

1. Copy this document
2. Update the data
3. Check it is "clean"
4. Make the same plot

It might seem repetitive, and there's often little sneaky bits in there, but it's just part of the process! Enjoy!

# Data Cleaning

```{r}
library(tidyverse)

education_2014 <- read_csv("raw_education_2014.csv")

education_2014

```

# Check age groups

Check to see how people have done the separations between ages

```{r}
table(education_2014$age_group)
```

We need to make sure that they are separated by `_` only:

```{r}
fixed_age_group_education_2014 <- education_2014 |>
  mutate(
    age_group = str_replace_all(
      age_group,
      pattern = "---|--|-",
      replacement = "_"
    ),
    age_group = as.factor(age_group)
  )
```


Check if this is fixed

```{r}
table(fixed_age_group_education_2014$age_group)
```

Check to see if year is right

```{r}
table(fixed_age_group_education_2014$year)
```

Good!

# Plot of proportion of people studying in each age group in each state

```{r}
# now let's explore the number of people studying in each age group in each state
ggplot(
  fixed_age_state_education_2014,
  aes(
    x = prop_studying,
    y = age_group
  )
) +
  geom_col() +
  facet_wrap(~state_territory, ncol = 2)

```

Why are there negative proportions?

```{r}
summary(fixed_age_state_education_2014$prop_studying)
hist(fixed_age_state_education_2014$prop_studying)
```


-99 means missing! Are there other instances of -99?


```{r}
library(naniar)

miss_scan_count(
  data = fixed_age_state_education_2014,
  search = -99
)
```

Replace -99 with NA

```{r}
tidy_age_state_education_2014 <- fixed_age_state_education_2014 |>
  mutate(
    prop_studying = na_if(prop_studying, -99)
  )

n_miss(tidy_age_state_education_2014)
```

Box plot, and a table that shows The 5 number summary (min, 1st quartile, median, 3rd quartile, max) of number educated for each state.

```{r}
ggplot(
  tidy_age_state_education_2014,
  aes(
    x = prop_studying,
    y = state_territory
  )
) +
  geom_boxplot()
```

# Table stating the summary of each state

```{r}
tidy_age_state_education_2014 |>
  group_by(state_territory) |>
  summarise(
    mean_study = mean(prop_studying),
    sd_study = sd(prop_studying)
  )
```

Drop NA values

```{r}
tidy_age_state_education_2014 |>
  group_by(state_territory) |>
  summarise(
    mean_study = mean(prop_studying, na.rm = TRUE),
    sd_study = sd(prop_studying, na.rm = TRUE)
  )
```

```{r}

education_2015 <- read_csv("raw_education_2015.csv")

education_2015
```
