---
title: "Solution Alpha: The Data Cleaning Document"
format: html
---

# Your Tasks

1. Write a function for each of the following:
  1. Bar plot of proportion of people educated by state and age
  2. Boxplot of **proportion of people educated** for each state
  3. Table of 5 number summary of **proportion of people educated** for each state
2. Move all these functions to the top of the document into a single code chunk labelled "functions"
3. Move the "data checks" to the end of the document. We will come back to this later
4. Make a "libraries" code chunk and put the library call in there


# Onboarding

Hi there! Welcome to the group! This is "The Data Cleaning Document". Here's the process

1. Read the data in
2. Go through various ad hoc tests
3. Make sure the data is "clean"
4. Produce a plot of the **proportion of people educated** in each age group in each state
5. Produce a box plot of **proportion of people educated** for each state.
6. Produce a table of The 5 number summary (min, 1st quantile, median, 3rd quantile, max)  of **proportion of people educated** for each state.

When we get another dataset, we make a new document to clean that data. The process is:

1. Copy this document
2. Update the data
3. Check it is "clean"
4. Make the same plots and tables

It might seem repetitive, and there's often little sneaky bits in there, but it's just part of the process! 

Enjoy!

# Data Cleaning

```{r}
#| label: read-data
library(tidyverse)

education_2014 <- read_csv("raw_education_2014.csv")

education_2014

```

## data quality checks

Check to see how people have done the separations between ages

```{r}
#| label: table-ages
table(education_2014$age_group)
table(education_2014$year)
```

Check summaries of data:

```{r}
#| label: summarise-prop-studying
summary(education_2014)
```

Maybe -99 means missing!? Are there other instances of -99?

```{r}
hist(education_2014$prop_studying)
```


```{r}
#| label: scan-missings
library(naniar)

miss_scan_count(
  data = education_2014,
  search = -99
)
```

## Functions for data cleaning

### clean age groups

```{r}
# We need to make sure that they are separated by `_` only:
clean_age_groups <- function(age_groups){
  clean_age <- str_replace_all(
      age_groups,
      pattern = "---|--|-",
      replacement = "_"
    )
  
  as.factor(clean_age)
}
```

### replace missing data

This actually probably doesn't need to have a function, since it is just one line of code.

### write a cleaning data function

```{r}
clean_education_data <- function(data){
  data |> 
  mutate(
    age_group = clean_age_groups(age_group),
    prop_studying = na_if(prop_studying, -99)
  )
}
```


```{r}
tidy_age_state_education_2014 <- clean_education_data(
  education_2014
)
```


# Produce a plot of the **proportion of people educated** in each age group in each state

```{r}
#| label: plot-prop-age
# now let's explore the number of people studying in each age group in each state
ggplot(
  tidy_age_state_education_2014,
  aes(
    x = prop_studying,
    y = age_group
  )
) +
  geom_col() +
  facet_wrap(~state_territory, ncol = 2)

```

# Produce a box plot of **proportion of people educated** for each state.

```{r}
#| label: boxplot-prop-studying
ggplot(
  tidy_age_state_education_2014,
  aes(
    x = prop_studying,
    y = state_territory
  )
) +
  geom_boxplot()
```

# Produce a table of The 5 number summary (min, 1st quantile, median, 3rd quantile, max)  of **proportion of people educated** for each state.

Drop NA values

```{r}
#| label: table-prop-studying-na
tidy_age_state_education_2014 |>
  group_by(state_territory) |>
  summarise(
    min_study = min(prop_studying, na.rm = TRUE),
    q1_studying = quantile(prop_studying, probs = 0.25, na.rm = TRUE),
    median_study = median(prop_studying, na.rm = TRUE),
    q3_studying = quantile(prop_studying, probs = 0.75, na.rm = TRUE),
    sd_study = sd(prop_studying, na.rm = TRUE)
  ) |> 
  knitr::kable(digits = 3)
```
