---
title: "Solution Alpha: The Data Cleaning Document"
format: html
---

# Your Tasks

1. Use `use_r()` from `devtools`/`usethis` to create a separate R file for each of the R functions:
  1. clean_age_groups
  2. clean_education_data
  3. plot_study_age_state
  4. boxplot_study_state
  5. summarise_prop_study
2. load all of these R functions, by calling `source` at the top of the quarto document

# Onboarding

Hi there! Welcome to the group! This is "The Data Cleaning Document". Here's the process

1. Read the data in
2. Go through various ad hoc tests
3. Make sure the data is "clean"
4. Produce a plot of the **proportion of people educated** in each age group in each state
5. Produce a box plot of **proportion of people educated** for each state.
6. Produce a table of The 5 number summary (min, 1st quantile, median, 3rd quantile, max)  of **proportion of people educated** for each state.

When we get another dataset, we make a new document to clean that data. The process is:

1. Copy this document
2. Update the data
3. Check it is "clean"
4. Make the same plots and tables

It might seem repetitive, and there's often little sneaky bits in there, but it's just part of the process! 

Enjoy!

# Libraries

```{r}
#| label: libraries
library(tidyverse)
library(naniar)
library(here)
```


# Functions

```{r}
#| label: functions

# clean age groups
# We need to make sure that they are separated by `_` only:
clean_age_groups <- function(age_groups){
  clean_age <- str_replace_all(
      age_groups,
      pattern = "---|--|-",
      replacement = "_"
    )
  
  as.factor(clean_age)
}

# write a cleaning data function

clean_education_data <- function(data){
  data |> 
  mutate(
    age_group = clean_age_groups(age_group),
    prop_studying = na_if(prop_studying, -99)
  )
}

# Functions for plotting and tables

plot_study_age_state <- function(data){
  ggplot(
  data = data,
  aes(
    x = prop_studying,
    y = age_group
  )
) +
  geom_col() +
  facet_wrap(~state_territory, ncol = 2)

}

boxplot_study_state <- function(data){
  ggplot(
  data,
  aes(
    x = prop_studying,
    y = state_territory
  )
) +
  geom_boxplot()

}

summarise_prop_study <- function(data){
  data |>
    group_by(state_territory) |>
    summarise(
      min_study = min(prop_studying, na.rm = TRUE),
      q1_studying = quantile(prop_studying, probs = 0.25, na.rm = TRUE),
      median_study = median(prop_studying, na.rm = TRUE),
      q3_studying = quantile(prop_studying, probs = 0.75, na.rm = TRUE),
      sd_study = sd(prop_studying, na.rm = TRUE)
  ) 
}

```

# Data Cleaning

```{r}
#| label: read-data
education_2014 <- read_csv(here("raw_education_2014.csv"))

education_2014

tidy_age_state_education_2014 <- clean_education_data(
  education_2014
)

tidy_age_state_education_2014
```



# Produce a plot of the **proportion of people educated** in each age group in each state

```{r}
#| label: plot-prop-age
plot_study_age_state(tidy_age_state_education_2014)

```

# Produce a box plot of **proportion of people educated** for each state.

```{r}
#| label: boxplot-prop-studying
boxplot_study_state(tidy_age_state_education_2014)
```

# Produce a table of The 5 number summary (min, 1st quantile, median, 3rd quantile, max)  of **proportion of people educated** for each state.

```{r}
#| label: table-prop-studying-na
summarise_prop_study(tidy_age_state_education_2014) |> 
  knitr::kable(digits = 3)
```

## data quality checks

Check to see how people have done the separations between ages

```{r}
#| label: table-ages
table(education_2014$age_group)
table(education_2014$year)
```

Check summaries of data:

```{r}
#| label: summarise-prop-studying
summary(education_2014)
```

Maybe -99 means missing!? Are there other instances of -99?

```{r}
hist(education_2014$prop_studying)
```


```{r}
#| label: scan-missings

miss_scan_count(
  data = education_2014,
  search = -99
)
```
