---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# learned

<!-- badges: start -->
<!-- badges: end -->

The goal of learned is to demonstrate some cleaning tools for education data.

## Installation

You can install the development version of learned from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("njtierney/learned")
```

## Example

Let's show an example of the raw education data:

```{r}
#| label: show-education-data

library(learned)

raw_education_2014

tidy_age_state_education_2014 <- clean_education_data(
  raw_education_2014
)

tidy_age_state_education_2014
```

### Plot of the **proportion of people educated** in each age group in each state

```{r}
#| label: plot-prop-age
plot_study_age_state(tidy_age_state_education_2014)
```

# Box plot of **proportion of people educated** for each state.

```{r}
#| label: boxplot-prop-studying
boxplot_study_state(tidy_age_state_education_2014)
```

# Table of The 5 number summary (min, 1st quantile, median, 3rd quantile, max)  of **proportion of people educated** for each state.

```{r}
#| label: table-prop-studying-na
summarise_prop_study(tidy_age_state_education_2014) |> 
  knitr::kable(digits = 3)
```
