---
title: "Solution Alpha: The Data Cleaning Document"
format: html
---

# Your Tasks

1. Move all the "data quality" checks into a new section called "data quality"
2. Move all of the data cleaning code up to the top, so we just work with one data set, named `tidy_age_state_education_2014`
3. Create two functions to clean the data: 
  1. tidy the age groups
  2. remove the missing values
4. Put these two functions into another function that does the data cleaning


# Onboarding

Hi there! Welcome to the group! This is "The Data Cleaning Document". Here's the process

1. Read the data in
2. Go through various ad hoc tests
3. Make sure the data is "clean"
4. Produce a plot of the **proportion of people educated** in each age group in each state
5. Produce a box plot of **proportion of people educated** for each state.
6. Produce a table of The 5 number summary (min, 1st quantile, median, 3rd quantile, max)  of **proportion of people educated** for each state.

When we get another dataset, we make a new document to clean that data. The process is:

1. Copy this document
2. Update the data
3. Check it is "clean"
4. Make the same plots and tables

It might seem repetitive, and there's often little sneaky bits in there, but it's just part of the process! 

Enjoy!

# Data Cleaning

```{r}
#| label: read-data
library(tidyverse)

education_2014 <- read_csv("raw_education_2014.csv")

education_2014

```

## data quality checks

Check to see how people have done the separations between ages

```{r}
#| label: table-ages
table(education_2014$age_group)
table(education_2014$year)
```

Check summaries of data:

```{r}
#| label: summarise-prop-studying
summary(education_2014)
```

Maybe -99 means missing!? Are there other instances of -99?

```{r}
hist(education_2014$prop_studying)
```


```{r}
#| label: scan-missings
library(naniar)

miss_scan_count(
  data = education_2014,
  search = -99
)
```

## Functions for data cleaning

### clean age groups

```{r}
# We need to make sure that they are separated by `_` only:
clean_age_groups <- function(age_groups){
  clean_age <- str_replace_all(
      age_groups,
      pattern = "---|--|-",
      replacement = "_"
    )
  
  as.factor(clean_age)
}
```

### replace missing data

This actually probably doesn't need to have a function, since it is just one line of code.

### write a cleaning data function

```{r}
clean_education_data <- function(data){
  data |> 
  mutate(
    age_group = clean_age_groups(age_group),
    prop_studying = na_if(prop_studying, -99)
  )
}
```


```{r}
tidy_age_state_education_2014 <- clean_education_data(
  education_2014
)
```


# Produce a plot of the **proportion of people educated** in each age group in each state

```{r}
#| label: plot-prop-age
# now let's explore the number of people studying in each age group in each state
ggplot(
  tidy_age_state_education_2014,
  aes(
    x = prop_studying,
    y = age_group
  )
) +
  geom_col() +
  facet_wrap(~state_territory, ncol = 2)

```

# Produce a box plot of **proportion of people educated** for each state.

```{r}
#| label: boxplot-prop-studying
ggplot(
  tidy_age_state_education_2014,
  aes(
    x = prop_studying,
    y = state_territory
  )
) +
  geom_boxplot()
```

# Produce a table of The 5 number summary (min, 1st quantile, median, 3rd quantile, max)  of **proportion of people educated** for each state.

Drop NA values

```{r}
#| label: table-prop-studying-na
tidy_age_state_education_2014 |>
  group_by(state_territory) |>
  summarise(
    min_study = min(prop_studying, na.rm = TRUE),
    q1_studying = quantile(prop_studying, probs = 0.25, na.rm = TRUE),
    median_study = median(prop_studying, na.rm = TRUE),
    q3_studying = quantile(prop_studying, probs = 0.75, na.rm = TRUE),
    sd_study = sd(prop_studying, na.rm = TRUE)
  ) |> 
  knitr::kable(digits = 3)
```
